---
title: "HW5"
author: "Maggie Trout"
date: "2025-02-12"
output: rmdformats::material
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Homework 5


# Problem 1 


Assign to the variable n_dims a single random integer between 3 and 10.
```{r}

n_dims <- sample(3:10,1)
print(n_dims)
```

Create a vector of consecutive integers from 1 to n_dims2
```{r}
vec_a <- seq(from=1, to=(n_dims^2))
print(vec_a)
```

Use the sample function to randomly reshuffle these values.
```{r}
vec_a2 <- sample(x=vec_a)
print(vec_a2)
```

create a square matrix with these elements.
```{r}
matrix_a <- matrix(data=vec_a2,nrow=n_dims)
print(matrix_a)
```

find a function in r to transpose the matrix (+ print it out again and note how it has changed)
```{r}
transposed_matrix_a <- t(matrix_a)
print(transposed_matrix_a)
```

calculate the sum and the mean of the elements in the first row and then the last row.

```{r}
sum(transposed_matrix_a[1, ])
sum(transposed_matrix_a[n_dims, ])
mean(transposed_matrix_a[1, ])
mean(transposed_matrix_a[n_dims, ])
```

read about the eigen() function and use it on your matrix
```{r}
decomp <- eigen(transposed_matrix_a)
```

look carefully at the elements of $values and $vectors in the output. What kind of numbers are these?
```
Both appear to be a number, either positive or negative plus or minus an imaginary number. However sometimes the imaginary number is 0i 
```

dig in with the typeof() function to figure out their type.

```{r}
typeof(decomp$values)
typeof(decomp$vectors)
```

if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.

To demonstrate:
```{r}
n_dims <- sample(3:10,1)
print(n_dims)

vec_a <- seq(from=1, to=(n_dims^2))
print(vec_a)

vec_a2 <- sample(x=vec_a)
print(vec_a2)

matrix_a <- matrix(data=vec_a2,nrow=n_dims)
print(matrix_a)

transposed_matrix_a <- t(matrix_a)
print(transposed_matrix_a)

sum(transposed_matrix_a[1, ])
sum(transposed_matrix_a[n_dims, ])
```

# Problem 2

Create a list with the following named elements:

my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
my_list <- list(my_matrix <- matrix(runif(16),nrow=4), 
                 my_logical <- runif(100)<0.5, 
                 my_letters <- sample(x=letters, size=26))
```

Then, complete the following steps:

create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
new_list <- list(my_matrix[[2,2]], my_logical[2], my_letters[2])
```
use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
```
combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
new_vec <- c(new_list[[1]],new_list[[2]],new_list[[3]])
print(new_vec)
```
what is the data type of this vector?
```{r}
typeof(new_vec)
```

# Problem 3 
Create a data frame with the two variables (= columns) and 26 cases (= rows) below:

call the first variable my_unis and fill it with 26 random uniform values from 0 to 10

call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
d_frame <- data.frame(my_unis=runif(n=26,min=0, max=10),
                      my_letters=sample(x=LETTERS, size=26))
```
for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
d_frame[sample(x=1:26,4),1] <- NA
```
for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
which(!complete.cases(d_frame))
```
re-order the entire data frame to arrange the second variable in alphabetical order
```{r}
d_frame <- d_frame[order(d_frame$my_letters), ]
```
calculate the column mean for the first variable.
```{r}
mean(d_frame$my_unis, na.rm = TRUE)
```
