---
title: "HW8"
author: "Maggie Trout"
date: "2025-03-19"
output: 
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Homework #8

Open libraries
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

This data is taken from calculation of the total number of mitochondrial branches in an image of a cell. It is a relatively small dataset (36 entries).

Load in data:
```{r}
z <- read.table("Mito_Branches.csv",header=TRUE,sep=",")
names(z) <- list("Branches") #naming the observations
str(z)
summary(z)
```

Plot histogram of data
```{r}
#plot histogram of the data
p1 <- ggplot(data=z, aes(x=Branches, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1) 
```
    
Add empirical density curve
```{r}
# Adding density curve of the data (not fitted)
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```


Get maximum likelihood parameters for normal
```{r}
normPars <- fitdistr(z$Branches,"normal") #fitting to normal distribution
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```
Plot normal probability density
```{r}
meanML <- normPars$estimate["mean"] # pulling out the mean and sd from the fit
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$Branches),len=length(z$Branches)) 

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$Branches), args = list(mean = meanML, sd = sdML))
 p1 + stat # adding the normal probability density to the plot
```

Plot exponential probability density 
```{r}
expoPars <- fitdistr(z$Branches,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$Branches), args = list(rate=rateML))
 p1 + stat + stat2
```

Plot uniform probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$Branches), args = list(min=min(z$Branches), max=max(z$Branches))) # uniform only needs the range of the data
 p1 + stat + stat2 + stat3
```
 
Plot gamma probability density
```{r}
gammaPars <- fitdistr(z$Branches,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$Branches), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

Plot beta probability density
```{r}
pSpecial <- ggplot(data=z, aes(x=Branches/(max(Branches + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$Branches/max(z$Branches + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$Branches), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial # beta needs a new graph
```

Looking at the graphs, the gamma distribution appears to fit the best (though not perfect- again this is a relatively small dataset)

# Simulating a new dataset based on best distribution:

Simulating data and making a histogram, with the gamma probability density 
```{r}
# simulate new data
set.seed(37) # set seed so simulated data remains the same when run
my_sim_data <- rgamma(n=36,shape=shapeML,rate=rateML)
my_sim_data <- data.frame(1:36,my_sim_data)
names(my_sim_data) <- list("ID","Branches")

# make a histogram of the new data
p2 <- ggplot(data=my_sim_data, aes(x=Branches, y=..density..)) +
  geom_histogram(color="grey60",fill="palegreen2",size=0.2) 

# add gamma probability density to the graph
gstat <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="magenta1", n = length(my_sim_data$Branches), args = list(shape=shapeML, rate=rateML))
 p2 + gstat 
```

Original data and gamma probability density:
```{r}
p1 <- ggplot(data=z, aes(x=Branches, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
 p1 + stat4
```

The simulated histogram is relatively similar to the real data, especially in that both have the majority of reads clumped between about 50 and 250, with a few reads appearing between 300-500. Obviously they are not identical, especially as this is a smaller dataset, but the overall shape of the histograms is pretty close. This is a good sign that the fit gamma distribution does represent the data relatively well. 
